---
title: |
  | Applied Machine Learning in Economics
  | Exercise 1
author: "Nasim Gholizadeh"
output: html_document
---

---------------------------------------------------------

# Question 1
Explain whether each scenario is a classification or regression problem, and indicate
whether we are most interested in inference or prediction. Finally, provide $n$ and $p$.

a) We collect a set of data on the top 500 firms in the US. For each firm we record
profit, number of employees, industry and the CEO salary. We are interested in
understanding which factors affect CEO salary.
b) We are considering launching a new product and wish to know whether it will be
a _success_ or a _failure_. We collect data on 20 similar products that were previously
launched. For each product we have recorded whether it was a success or failure,
price charged for the product, marketing budget, competition price, and ten other
variables.
c) We are interested in predicting the % change in the USD/Euro exchange rate in
relation to the weekly changes in the world stock markets. Hence we collect weekly
data for all of 2012. For each week we record the % change in the USD/Euro, the
% change in the US market, the % change in the British market, and the % change
in the German market.

# Answer 1
a) Regression problem as the response variable is quantitative.
We are interested in inference because we are interested in understanding which factors affect CEO salary.
n : 500 (500 firms). n=500
p : 3 (profit, number of employees, and industry). p = 3

b) Classification problem since the response variable is a binary value (success or failure).
We are interested in prediction, whether the new product will be a success or a failure.
n : 20 (20 similar products). n=20
p : 13 (price charged for the product, marketing budget, competition price, and 10 other variables). p = 13

c) Regression problem because the response variable (% change in the USD/Euro) is a quantitative value.
We are interested in prediction of the % change in the USD/Euro exchange rate in relation to the weekly changes in the world stock markets.
n : 52 (weekly data for all of 2012). n=52
p : 3 (the % change in the US market, the % change in the British market, and the % change in the German market). p = 3


# Question 2 (applied)
This exercise relates to the `College` data set, It contains a number of variables for 777 different universities and colleges in the US. The variables are

- `Private`: Public/private indicator
- `Apps`: Number of applications received
- `Accept`: Number of applicants accepted
- `Enroll`: Number of new students enrolled
- `Top10perc`: New students from top 10
- `Top25perc`: New students from top 25
- `F.Undergrad`: Number of full-time undergraduates
- `P.Undergrad`: Number of part-time undergraduates
- `Outstate`: Out-of-state tuition
- `Room.Board`: Room and board costs
- `Books`: Estimated book costs
- `Personal`: Estimated personal spending
- `PhD`: Percent of faculty with Ph.D.'s
- `Terminal`: Percent of faculty with terminal degree
- `S.F.Ratio`: Student/faculty ratio
- `perc.alumni`: Percent of alumni who donate
- `Expend`: Instructional expenditure per student
- `Grad.Rate`: Graduation rate

Before reading the data into R, it can be viewed in Excel or a text editor.

a) Use the `read.csv()` function to read the data into `R`. Call the loaded data `college`. Make sure that you have the directory set to the correct location for the data. The `R` commands `getwd()` and `setwd()` may be helpful.

b) Look at the data using the `head()` function. You should notice that the first column is just the name of each university. We don't really want R to treat this as data. However, it may be handy to have these names for later. Try the following commands:

```{r, eval=FALSE}
rownames(college) <- college[, 1]
View(college)
```

You should see that there is now a `row.names` column with the name of each university recorded. This means that `R` has given each row a name corresponding to the appropriate university. `R` will not try to perform calculations on the row names. However, we still need to eliminate the first column in the data where the names are stored. Try

```{r, eval=FALSE}
college <- college[, -1]
View(college)
```

Now you should see that the first data column is `Private`. Note that another column labeled `row.names` now appears before the `Private` column. However, this is not a data column but rather the name that `R` is giving to each row.

c) 
i. Use the `summary()` function to produce a numerical summary of the variables in
the data set.
ii. Use the `pairs()` function to produce a scatterplot matrix of the first ten columns
or variables of the data. Recall that you can reference the first ten columns of a
matrix `A` using `A[,1:10]`.
iii. Use the `plot()` function to produce side-by-side boxplots of `Outstate` versus
`Private`.
iv. Create a new qualitative variable, called `Elite`, by binning the `Top10perc`
variable. We are going to divide universities into two groups based on whether or
not the proportion of students coming from the top 10% of their high school classes
exceeds 50%.


```{r, eval=FALSE}
Elite <- rep("No", nrow(college))
Elite[college$Top10perc > 50] <- "Yes"
Elite <- as.factor(Elite)
college <- data.frame(college, Elite)
```

Use the `summary()` function to see how many elite universities there are. Now use the `plot()` function to produce side-by-side boxplots of `Outstate` versus `Elite`.

v. Use the `hist()` function to produce some histograms with differing numbers of
bins for a few of the quantitative variables. You may find the command `par(mfrow
= c(2, 2))` useful: it will divide the print window into four regions so that four
plots can be made simultaneously. Modifying the arguments to this function will
divide the screen in other ways.

vi. Continue exploring the data, and **provide a brief summary** of what you discover.

# Answer 2 (applied)

a)
```{r}
college <- read.csv("College.csv", stringsAsFactors=T)
```

b)
```{r}
rownames(college) <- college[, 1]
college <- college[, -1]
head(college)
```

c)
i)
```{r }
summary(college)
```
ii)
```{r }
pairs(college[, 1:10])
```
iii)
```{r }
plot(college$Private, college$Outstate, main = "Outstate versus Private", 
    xlab = "Private", ylab = "Outstate")
```

iv)
```{r }
Elite <- rep("No", nrow(college))
Elite[college$Top10perc > 50] <- "Yes"
Elite <- as.factor(Elite)
college <- data.frame(college, Elite)
summary(college$Elite)
```

```{r }
plot(college$Elite, college$Outstate, main = "Outstate versus Elite", 
    xlab = "Elite", ylab = "Outstate")
```

## Similarly
# boxplot(college$Outstate ~ college$Elite, main = "Outstate versus Elite", 
#     xlab = "Elite", ylab = "Outstate")

v)
```{r }
par(mfrow = c(3,3))
hist(college$Accept, main = "Histogram of Accept", xlab = "Accept")
hist(college$Enroll, main = "Histogram of Enroll", xlab = "Enroll")
hist(college$Top10perc, main = "Histogram of Top10perc", xlab = "Top10perc")
hist(college$F.Undergrad, main = "Histogram of F.Undergrad", xlab = "F.Undergrad")
hist(college$Outstate, main = "Histogram of Outstate", xlab = "Outstate")
hist(college$Books, main = "Histogram of Books", xlab = "Books")
hist(college$PhD, main = "Histogram of PhD", xlab = "PhD")
hist(college$Expend, main = "Histogram of Expend", xlab = "Expend")
hist(college$Grad.Rate, main = "Histogram of Grad.Rate", xlab = "Grad.Rate")

```
vi)
```{r}
ols.fit <- lm(Grad.Rate ~ Apps + Top25perc + Personal + F.Undergrad + perc.alumni + Expend, 
            data = college)
summary(ols.fit)
```
